---
title: "STAA 566 - Homework 4 - Table"
author: "Alexis Suhr"
date: "10 March 2022"
output: html_document
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---



### Assignment Summary   
**Make a dynamic figure. Describe the following:**  


**1. Data Source(s)**    


(a) 'mpg' from the 'ggplot2' package**\
(38 observations of popular car models' make, class, engine size, transmission, and fuel economy in 1999 and 2008)

(b) 'Transportation Energy-Related Carbon Dioxide Emissions' from the U.S. Energy Information Administration (EIA) website** *(CO2 emissions by U.S. state for the years 1980 - 2018)


**2. What I want to convey in the Table(s)**  
I want to show users the CO2 emmissions data in the same context as the 'mpg' dataset from 'ggplot2' to give a related idea of emissions with other vehicular transportation statistics.  

**3. What functionality I put in the Table(s)/What formatting I put in the Table(s) and why **  
I wanted to make the tables easy to read and concise. I added the 'hover' function as a way for readers to highlight the portion of the table they're reading. I also changed style options including font type, font size, and was able to put the title in bold and with larger font than the rest of its corresponding table.  


### Libraries, Packages

```{r include=FALSE, warning = FALSE}
###### 
# Read in libraries to be used 
###### 
# install.packages(c('ggplot2',
#           'DataExplorer',
#           'tidyverse',
#           'plotly',
#           'readr',
#           'rworldmap',
#           'sf',
#           'rgeos',
#           'poliscidata',
#           'tigris',
#           'usmap',
#           'viridis',
#           'gridExtra',
#           'shiny',
#           'shinydashboard',
#           'babynames',
#           'kableExtra',
#           'R.utils',
#           'summarytools'))

libs <- c('ggplot2',
          'DataExplorer',
          'tidyverse',
          'plotly',
          'readr',
          'rworldmap',
          'sf',
          'rgeos',
          'poliscidata',
          'tigris',
          'usmap',
          'viridis',
          'gridExtra',
          'shiny',
          'shinydashboard',
          'babynames',
          'kableExtra',
          'R.utils',
          'summarytools',
          'readr')


for(l in libs){
  suppressPackageStartupMessages(library(l,
                                         quietly = TRUE,
                                         character.only = TRUE))
}
```

### Reading in Data to Be Used for Analysis  

```{r echo = FALSE, warning = FALSE}
#####
# Loading in 'babynames' to verify shiny dashboard 
# creation/commands used in relation to the dataset
#####
# **Do not need to include in final knitting process**
# **COMMENT OUT**
# babynames_ExampleData <- babynames::babynames

#####
# Fuel Economy data from 'mpg' in 'ggplot2'
#####
# ?mpg
FuelEconData <- ggplot2::mpg %>%
  mutate(Manufacturer = as.factor(R.utils::capitalize(manufacturer)), 
         
         # Name of manufacturer
         Model = as.factor(toupper(model)),
         # Model of car
         
         EngineDisplacement = displ, 
         # Liters of engine displacement (total volume of all the 
         # cylinders in an engine)
         
         YearManufactured = as.factor(year), 
         # Year manufactured
         
         NumCylinders = as.factor(cyl), 
         # Number of engine cylinders (4, 5, 6, or 8)
         
         TransmissionType1 = as.factor(substr(R.utils::capitalize(trans), 1, nchar(trans)-4)),
         TransmissionType2 = as.factor(toupper(substr(trans, nchar(trans)-2, nchar(trans)))),
         # The above two variables are separating the 'trans' string 
         # and capitalizing certain values to make the data look nicer
         
         TransmissionType = as.factor(paste0(TransmissionType1, " (", TransmissionType2)),
         # Type of engine transmission
         
         DriveTrain = as.factor(case_when(drv =='f' ~ 'Front-Wheel',
                                          drv == 'r' ~ 'Rear-Wheel',
                                          drv == '4' ~ 'Four-Wheel')), 
         # Type of Drive Train 
         #(front-wheel, rear-wheel, 4-wheel)
         
         CityMPG = cty, 
         # City fuel efficiency (mpg)
         
         HighwayMPG = hwy, 
         # Highway fuel efficiency (mpg)
         
         FuelType = as.factor(case_when(fl == 'e' ~ 'Ethanol-85',
                                        fl == 'd' ~ 'Diesel',
                                        fl == 'r' ~ 'Regular',
                                        fl == 'p' ~ 'Premium',
                                        fl == 'c' ~'Methane (CNG)')), 
         # Type of fuel used to run vehicle 
         # (ethanol 85,diesel,regular,premium,compressed natural gas aka Methane/CNG)
         CarType = as.factor(case_when(class == '2seater' ~ 'Two-Seater',
                                       class == 'compact' ~ 'Compact',
                                       class == 'midsize' ~ 'Midsize',
                                       class == 'minivan' ~ 'Minivan',
                                       class == 'pickup' ~ 'Pickup',
                                       class == 'subcompact' ~ 'Subcompact',
                                       class == 'suv' ~ 'SUV'))
         # Type of car build 
         # (2seater, compact, midsize, minivan, pickup, subcompact, suv)
  ) %>% 
  select(Manufacturer:CarType, 
         -c(TransmissionType1, 
            TransmissionType2)) %>%
  filter(CarType != 'Two-Seater',
         NumCylinders != '5')

#####
# CO2 emissions and Changes in CO2 emisions for 1999, 2008, 2015
#####

CO2EmissionsData <- read.csv("~/HW4_Suhr_Alexis/Datasets/CO2Emissions.csv") %>%
  mutate(state = state.abb[match(ï..State, state.name)],
         CO2_1999 = X1999,
         CO2_2008 = X2008,
         Change_1999_2008 = CO2_2008 - CO2_1999) %>%
  select(state,
         CO2_1999,
         CO2_2008,
         Change_1999_2008)


#####
# Viewing first 6 rows of each dataset
#####

# kable(head(FuelEconData))
# head(CO2EmissionsData)
```

### Tables  


```{r fig.align = "center", eval = TRUE, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE}

###### 
# 'mpg' Data from the 'ggplot2' Package  
#####

SummaryTable_1999 <- summarytools::descr(FuelEconData %>%
                             select(YearManufactured,
                                    CityMPG, 
                                    HighwayMPG) %>%
                             filter(YearManufactured == '1999'))

SummaryTable_2008 <- summarytools::descr(FuelEconData %>%
                             select(YearManufactured,
                                    CityMPG, 
                                    HighwayMPG) %>%
                             filter(YearManufactured == '2008'))


Kable_1999Summary <- SummaryTable_1999[c(1:7, 9), ] %>%
  kbl(caption = "<b>Fuel Efficiency (1999)<b>",
      digits = 2,
      col.names = c("City MPG",
                    "Highway MPG"),
        align = "c") %>%
    kable_minimal() %>%
    kable_paper("hover",
                full_width = F,
                html_font = "Times",
                font_size = 18) %>%
  gsub("font-size: initial !important;",
       "font-size: 20pt !important;",
       .) 


Kable_2008Summary <- SummaryTable_2008[c(1:7, 9), ] %>%
  kbl(caption = "<b>Fuel Efficiency (2008)<b>",
      digits = 2,
      col.names = c("City MPG",
                    "Highway MPG"),
        align = "c") %>%
    kable_minimal() %>%
    kable_paper("hover",
                full_width = F,
                html_font = "Times",
                font_size = 18) %>%
  gsub("font-size: initial !important;",
       "font-size: 20pt !important;",
       .)

#### 'CO2 Emissions' Data from the United States EIA   

SummaryTable_CO2 <- descr(CO2EmissionsData[c(2:4), ])


Kable_CO2Summary <- SummaryTable_CO2[c(1:7, 9), ] %>%
    kbl(caption = "<b>CO2 Emmissions by Year<b>",
        digits = 2,
        col.names = c("1999",
                      "2008",
                      "(2008-1999)"),
        align = "c") %>%
    kable_minimal() %>%
    kable_paper("striped", 
                "hover",
                full_width = F,
                html_font = "Times",
                font_size = 18) %>%
    footnote(symbol=c("Note: (2008-1999) is the difference in CO2 emissions from the year 1999 to the year 2008, obtained by subtracting column 1999 from column 2008")) %>%
  gsub("font-size: initial !important;",
       "font-size: 20pt !important;",
       .) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))



```

```{r echo = FALSE, warning = FALSE}
#####
# Summary Tables
#####
Kable_1999Summary
Kable_2008Summary
Kable_CO2Summary

```
